How fluid-mechanical erosion creates anisotropic porous media

### 2022-06-29 docs/PhysicsPaper/StokesErosion.tex
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quadrature for Cauchy integrals}
The accuracy of our method is determined by the accuracy of the
quadrature rule applied to equation~\eqref{eqn:FredBIE}. Since we have
written the Stokes double-layer potential
velocity~\eqref{eqn:velocityCauchy}, deformation
tensor~\eqref{eqn:deformationCauchy}, and
vorticity~\eqref{eqn:vorticityCauchy} as a sum of Cauchy integrals and
their derivatives, the accuracy of the method depends on the accuracy of
computing a general Cauchy integral
\begin{align}
  v(x) = \frac{1}{2\pi i} \int_{\gamma} \frac{\eta(y)}{x-y} \, dy.
  \label{eqn:cauchy1}
\end{align}
Here we describe a quadrature formulae that was first used to
approximate analytic functions~\cite{ioa-pap-per1991}, and then extended
to Stokes layer potentials~\cite{bar-wu-vee2015}. The quadrature method
requires the boundary data of the analytic function~\eqref{eqn:cauchy1}
which satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  v^{-}(x_0) = \lim_{\substack{x \rightarrow x_0 \\ x \in \Omega}} \int_{\gamma}
    \frac{\eta(y)}{x-y}\, dy = -\frac{1}{2}\eta(x_0) - 
    \frac{1}{2\pi i} \int_{\gamma} \frac{\eta(y)}{x_0-y} \, dy,
    \quad x_0 \in \gamma,
    \label{eqn:SP}
\end{align}
where the last integral is interpreted in the principal-value sense.
Here we are assuming that $\Omega$ is the bounded region interior to
$\gamma$. Once $v^{-}$ is calculated, $v(x)$ and its derivatives can be
determined by its boundary data alone using the Cauchy Integral Theorem
\begin{subequations}
  \label{eqn:cauchy}
  \begin{alignat}{3}
  \label{eqn:cauchyv}
  v(x) &= \frac{1}{2\pi i}\int_{\gamma} 
    \frac{v^{-}(y)}{y-x} \,dy, \\
  v'(x) &= \frac{1}{2\pi i} \int_{\gamma}
    \frac{v^{-}(y)}{(y-x)^2} \, dy, \\
  v''(x) &= \frac{1}{\pi i} \int_{\gamma}
    \frac{v^{-}(y)}{(y-x)^3} \, dy.
  \end{alignat}
\end{subequations}

@@

The previous section shows intriguing features to arise from the erosion of a random initialization of solid bodies. The observations natural raise the question: are the trends specific to the particular simulation shown or do these features emerge robustly across different initial conditions? To answer this question, we now analyze the same quantities---permeability, tortuosity, and anisotropy---over an ensemble of simulations having different numbers of bodies and different random initializations. We analyze runs having $M$ = 20, 40, 60, 80, and 100 bodies, with at least three different instances of each. This data set represents roughly 2,100 hours of computational time.

@@

 	As before, the anisotropy can be decomposed into configuration and shape components, $\anis_C$ and $\anis_S$ respectively, and the statistics of these quantities are shown in Figs.~\ref{fig5}(d)--(e). The shape anisotropy, seen in Fig.~\ref{fig5}(e), takes a fairly regular, parabolic arc --- first increasing due to the shapes carved by erosion and then decreasing as the bodies vanish. This behavior is consistent across all of the values of $M$. The configurational anisotropy (Fig.~\ref{fig5}(d)), however, shows less regular behavior. For runs with a smaller number of bodies ($M$ = 20 and 40) the configurational anisotropy remains relatively small throughout the entire simulation, indicating that the majority of observed anisotropy is due to the individual shapes of bodies. For the runs with a higher body count ($M $ = 60, 80, and 100), though, the configurational anisotropy grows moderately and then surges late in the simulation, as was observed in the single 80-body simulation from Fig.~\ref{fig3}. Since the shape anisotropy is decreasing at this time, the observed second rise in the total anisotropy is due entirely to this surge in configurational component. For these high-body count runs, the latest stages of erosion are dominated by channelization, which substantially promotes the configurational anisotropy over shape anisotropy. That is, the large-scale arrangement of the bodies has greater effect than the shape of individual bodies at this late stage.

@@

As before, the tortuosity ratio, $(T_2 -1)/(T_1 -1)$, provides a second proxy for medium anisotropy to complement the permeability-based definition. Figure~\ref{fig6}(c) shows that the tortuosity ratio behaves similarly across all simulations and all different values of $M$. The ratio increases substantially with erosion, reaching a peak of roughly $(T_2 -1)/(T_1 -1) \approx 8$, before descending during the final stages of erosion. As before the tortuosity ratio can be decomposed into configurational and shape components, as shown in Figures~\ref{fig6}(d)--(e). Here, we see that the shape component is the main contribution and the configurational component remains minimal. This observation fits with the intuition that the main hindrance to a passive tracer should be the shape of individual obstructions, rather than how those obstructions are arranged relative to one another. As such, the tortuosity ratio is a less sensitive indicator of channelization seen to develop late in the erosion simulations.

@@

In this paper, we have presented a Cauchy formulation of the boundary integral equation method to simulate the fluid-mechanical erosion of many bodies in a Stokes flow. The accuracy and efficiency of the method enables high-fidelity simulations of dense suspensions of $O(100)$ bodies and statistical analysis across an ensemble of different initializations. By linking the governing Stokes equations to Darcy's law, we are able to extract porous-media properties, including permeability, resistivity, tortuosity, and anisotropy, as they evolve over time. Direct comparison between the resistivity and the total drag force confirms such bulk properties are extracted with high accuracy.

The ratio of longitudinal to transverse permeability provides our main diagnostic of medium anisotropy. Measurements indicate that the anisotropy grows substantially as a result of fluid-mechanical erosion, reaching a peak of roughly six in most simulations, before the bodies completely vanish and return the system to an isotropic state. The anisotropy can be further decomposed into a configurational component, due only to the relative positions of bodies, and a shape component, due to the detailed geometry of each body. Statistical analysis across a large number of simulations reveals that if the number of bodies is sufficiently large, $M \ge 60$, the configurational anisotropy surges near the final stages of erosion and surpasses the shape anisotropy as the primary contribution. This surge in configurational anisotropy is associated with the formation of visual channels that transmit a large portion of the flow. The tortuosity provides a second metric and also reveals strong anisotropy to develop across all simulations. The methodology could be extended to more complex scenarios of thermal convective flows in porous media~\cite{McCurdy2019}, erodable porous structures formed by precipitation reactions~\cite{eastham2020multiphase}, and first arrival statistics of Brownian particles~\cite{che-lin-her-qua2021}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quadrature for Cauchy integrals}
The accuracy of our method is determined by the accuracy of the
quadrature rule applied to equation~\eqref{eqn:FredBIE}. Since we have
written the Stokes double-layer potential
velocity~\eqref{eqn:velocityCauchy}, deformation
tensor~\eqref{eqn:deformationCauchy}, and
vorticity~\eqref{eqn:vorticityCauchy} as a sum of Cauchy integrals and
their derivatives, the overall accuracy hinges on the computation of a general Cauchy integral
\begin{align}
  v(x) = \frac{1}{2\pi i} \int_{\gamma} \frac{\eta(y)}{x-y} \, dy.
  \label{eqn:cauchy1}
\end{align}
Here we describe a quadrature formulae that was first used to
approximate analytic functions~\cite{ioa-pap-per1991}, and then extended
to Stokes layer potentials~\cite{bar-wu-vee2015}. The quadrature method
requires the boundary data of the analytic function~\eqref{eqn:cauchy1}
which satisfies the Sokhotski-Plemelj jump relation
\begin{align}
  v^{-}(x_0) = \lim_{\substack{x \rightarrow x_0 \\ x \in \Omega}} \int_{\gamma}
    \frac{\eta(y)}{x-y}\, dy = -\frac{1}{2}\eta(x_0) - 
    \frac{1}{2\pi i} \int_{\gamma} \frac{\eta(y)}{x_0-y} \, dy,
    \quad x_0 \in \gamma,
    \label{eqn:SP}
\end{align}
where the last integral is interpreted in the principal-value sense.
Here we are assuming that $\Omega$ is the bounded region interior to
$\gamma$. Once $v^{-}$ is calculated, $v(x)$ and its derivatives can be
determined by its boundary data alone using the Cauchy Integral Theorem
\begin{subequations}
  \label{eqn:cauchy}
  \begin{alignat}{3}
  \label{eqn:cauchyv}
  v(x) &= \frac{1}{2\pi i}\int_{\gamma} 
    \frac{v^{-}(y)}{y-x} \,dy, \\
  v'(x) &= \frac{1}{2\pi i} \int_{\gamma}
    \frac{v^{-}(y)}{(y-x)^2} \, dy, \\
  v''(x) &= \frac{1}{\pi i} \int_{\gamma}
    \frac{v^{-}(y)}{(y-x)^3} \, dy.
  \end{alignat}
\end{subequations}

@@

The previous section shows intriguing features to arise from the erosion of a single, random initialization of solid bodies. The observations natural raise the question: are the trends specific to the particular simulation shown or do these features emerge robustly across different initial conditions? To answer this question, we now analyze the same quantities---permeability, tortuosity, and anisotropy---over an ensemble of simulations having different numbers of bodies and different random initializations. We analyze runs having $M$ = 20, 40, 60, 80, and 100 bodies, with at least three different instances of each. This data set represents roughly 2,100 hours of computational time.

@@

 	As before, the anisotropy can be decomposed into configuration and shape components, $\anis_C$ and $\anis_S$ respectively, and the statistics of these quantities are shown in Figs.~\ref{fig5}(d)--(e). The shape anisotropy, seen in Fig.~\ref{fig5}(e), takes a fairly regular, parabolic arc --- first increasing due to the shapes carved by erosion and then decreasing as the bodies vanish. This behavior is consistent across all of the values of $M$. The configurational anisotropy (Fig.~\ref{fig5}(d)), however, shows less regular behavior. For runs with a smaller number of bodies ($M$ = 20 and 40) the configurational anisotropy remains relatively small throughout the entire simulation, indicating that the majority of observed anisotropy is due to the individual shapes of bodies. For the runs with a higher body count ($M $ = 60, 80, and 100), though, the configurational anisotropy grows moderately and then surges late in the simulation, as was observed in the single 80-body simulation from Fig.~\ref{fig3}. Since the shape anisotropy is decreasing at this time, the observed second rise in the total anisotropy is due entirely to this surge in the configurational component. For these high-body count runs, the latest stages of erosion are dominated by channelization, which substantially promotes the configurational anisotropy over shape anisotropy. That is, the large-scale arrangement of the bodies has greater effect than the shape of individual bodies.

@@

As before, the tortuosity ratio, $(T_2 -1)/(T_1 -1)$, provides a second proxy for medium anisotropy to complement the permeability-based definition. Figure~\ref{fig6}(c) shows that the tortuosity ratio behaves similarly across all simulations and all different values of $M$. The ratio increases substantially with erosion, reaching a peak of roughly $(T_2 -1)/(T_1 -1) \approx 8$, before descending during the final stages of erosion. As before the tortuosity ratio can be decomposed into configurational and shape components, as shown in Figures~\ref{fig6}(d)--(e). Here, we see that the shape component is the main contribution and the configurational component remains minimal. This observation fits with the intuition that the main hindrance to a passive tracer should be the shape of individual obstructions, rather than how those obstructions are arranged relative to one another. As such, the tortuosity ratio is a less sensitive indicator of the channelization seen to develop late in the erosion simulations.

@@

In this paper, we have presented a Cauchy formulation of the boundary integral equations to simulate the fluid-mechanical erosion of many bodies in a Stokes flow. The accuracy and efficiency of the method enables high-fidelity simulations of dense suspensions of $O(100)$ bodies and statistical analysis across an ensemble of different initializations. By linking the governing Stokes equations to Darcy's law, we are able to extract porous-media properties, including permeability, resistivity, tortuosity, and anisotropy, as they evolve over time. Direct comparison between the resistivity and the total drag force confirms such bulk properties are extracted with high accuracy.

The ratio of longitudinal to transverse permeability provides our main diagnostic of medium anisotropy. Measurements indicate that the anisotropy grows substantially as a result of fluid-mechanical erosion, reaching a peak of roughly six in most simulations, before the bodies completely vanish and return the system to an isotropic state. The anisotropy can be further decomposed into a configurational component, due only to the relative positions of bodies, and a shape component, due to the detailed geometry of each body. Statistical analysis across a large number of simulations reveals that if the number of bodies is sufficiently large, $M \ge 60$, the configurational anisotropy surges near the final stages of erosion and surpasses the shape anisotropy as the primary contribution. This surge in configurational anisotropy is associated with the formation of visual channels that transmit a large portion of the flow. The tortuosity provides a second metric and also reveals strong anisotropy to develop across all simulations. In the future, we hope to extend the methodology to more complex scenarios of thermal convective flows in porous media~\cite{McCurdy2019}, erodable porous structures formed by precipitation reactions~\cite{eastham2020multiphase}, and first arrival statistics of Brownian particles~\cite{che-lin-her-qua2021}.
